---
title: "Useful summary calculations"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

tar_load(pcr_all)
tar_load(pcr_pos)
tar_load(coinf_df)
tar_load(pos_final)
tar_load(model_data_all)
tar_load(model_list)

tar_load(binom_tests_vir_fam)
tar_load(binom_tests_vir_fam_bats)
tar_load(binom_tests_vir_fam_rodents)

library(tidyverse)
library(knitr)
library(kableExtra)
```

## Results

```{r data subsets, echo=TRUE}

# unique specimens
unique_specimens <- pcr_all %>% 
  distinct(specimen_id, .keep_all = TRUE) 

# unique animals
unique_animals <- pcr_all %>% 
  distinct(predict_sample_id, .keep_all = TRUE) 
```

```{r summary values, echo=TRUE}
# number of unique specimens
n_unique_specimens <- nrow(unique_specimens)
n_unique_specimens

# number of unique specimens by country
spec_by_country <- unique_specimens %>% 
  group_by(country) %>% 
  dplyr::summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  mutate(perc = n/n_unique_specimens*100)
spec_by_country

# countries with 5k + specimens
countries_5k <- spec_by_country %>% 
  dplyr::filter(n > 5000) %>% 
  select(country)
countries_5k

# number of unique specimens by host order
spec_by_order <- unique_specimens %>% 
  group_by(host_order) %>% 
  dplyr::summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  mutate(perc = round(n/n_unique_specimens*100, 1))
spec_by_order

# number of PCR tests performed
n_tests <- nrow(pcr_all)
n_tests

# number of PCR tests by virus tested for
table(pcr_all$virus_target_tested)
  
# number of unique animals
n_unique_animals <- nrow(unique_animals)
n_unique_animals

# number of unique animals by order
table(unique_animals$host_order)

# number of animals infected with one or more viruses
n_inf_animals <- nrow(coinf_df)
n_inf_animals

# overall prevalence (number animals pos. for 1+ virus/number of unique animals)
overall_prev <- round(n_inf_animals/n_unique_animals*100, 2)
overall_prev

# infected animals by country
coinf_df %>% 
  group_by(country) %>% 
  dplyr::summarise(n = n()) %>% 
  arrange(desc(n)) 

# number of coinfected animals
n_coinf_animals <- coinf_df %>% 
  dplyr::filter(n_virus > 1) %>% 
  nrow()
n_coinf_animals

# frequency of single/double/triple infection
coinf_freq <- table(coinf_df$n_virus)

# as percentages
coinf_perc <- round(coinf_freq/n_inf_animals*100,2)

# within COINFECTED animals, what was the breakdown by host order?
coinf_order_summary <- coinf_df %>% 
  dplyr::filter(n_virus > 1) %>% 
  group_by(host_order) %>% 
  summarize(n = n()) %>% 
  mutate(perc = round(n / sum(n)*100, 1))
coinf_order_summary

# within ALL infected animals, how common is coinfection by host order?
coinf_within_inf <- xtabs(~coinf_df$host_order + coinf_df$virus_coinf) %>% 
  as.data.frame.matrix() %>% 
  rename(single_inf = "0", coinf = "1") %>% 
  mutate(n_tot = single_inf + coinf,
         prev = round(coinf/n_tot*100, 2))
coinf_within_inf

# prevalence by host order (within unique animals) for orders with at least one infected animal
order_prev <- table(coinf_df$host_order) %>% 
  data.frame() %>% 
  rename(host_order = Var1, n_inf = Freq) %>% 
  left_join(
    table(unique_animals$host_order) %>% 
      data.frame() %>% 
      rename(host_order = Var1, n_tot = Freq)) %>% 
  mutate(prev = round(n_inf/n_tot*100, 2))
order_prev

# coinfection by host order within the model data (subset of animals with age and sex)
model_coinf_within_inf <- xtabs(~host_order + virus_coinf, model_data_all) %>% 
  as.data.frame.matrix() %>% 
  rename(single_inf = "0", coinf = "1") %>% 
  mutate(n_tot = single_inf + coinf,
         prev = round(coinf/n_tot*100, 2))
model_coinf_within_inf

```

### Methods (Overview of cleaned data)
* Specimen collection (n = `r n_unique_specimens` unique specimens from `r n_unique_animals` unique animals) was concentrated in countries in West and Central Africa as well as South and Southeast Asia, with > 5,000 specimens collected in `r countries_5k[1, 1]`, `r countries_5k[2, 1]`, `r countries_5k[3, 1]`, `r countries_5k[4, 1]`, `r countries_5k[5, 1]`, `r countries_5k[6, 1]`, and `r countries_5k[7, 1]` each (Fig. S1). 
* Most specimens were collected from bats (order Chiroptera; n = `r spec_by_order[1,2]`, `r spec_by_order[1,3]`%) and rodents (order Rodentia; n = `r spec_by_order[2,2]`, `r spec_by_order[2,3]`%). 
* There were `r n_tests` total PCR tests.

### Results (overview)
* Out of `r n_unique_animals` unique animals, `r n_inf_animals` (`r overall_prev`%) were infected with one or more viruses (Fig. 1A).
* Viral prevalence was ≥ 20% in Passeriformes (`r order_prev[6,2]`/`r order_prev[6,3]`; `r order_prev[6,4]`%), Anseriformes (`r order_prev[1,2]`/`r order_prev[1,3]`; `r order_prev[1,4]`%), and Galliformes (`r order_prev[5,2]`/`r order_prev[5,3]`; `r order_prev[5,4]`%).
* Prevalence in Chiroptera (`r order_prev[3,2]`/`r order_prev[3,3]`; `r order_prev[3,4]`%) was somewhat higher than average, while prevalence in Rodentia (`r order_prev[7,2]`/`r order_prev[7,3]`; `r order_prev[7,4]`%) and Eulipotyphla (`r order_prev[4,2]`/`r order_prev[4,3]`; `r order_prev[4,4]`%) was slightly lower than average (Table S1).
* A single virus was detected in most infected animals (n = `r coinf_freq[1]`, `r coinf_perc[1]`%); coinfections of two (n = `r coinf_freq[2]`, `r coinf_perc[2]`%) or three (n = `r coinf_freq[3]`, `r coinf_perc[3]`%) viruses in the same animal were infrequent. Of the `r n_coinf_animals` animals in which virus coinfection was detected, over two-thirds were Chiroptera (n = `r coinf_order_summary[2,2]`, `r coinf_order_summary[2,3]`%) with the remainder being `r coinf_order_summary[4,2]` Rodentia, `r coinf_order_summary[1,2]` Anseriformes, and `r coinf_order_summary[3,2]` Eulipotyphla. Within infected animals, virus coinfection rates were highest in Anseriformes (`r coinf_within_inf[1, 2]`/`r coinf_within_inf[1, 3]`, `r coinf_within_inf[1, 4]`%), Rodentia (`r coinf_within_inf[7, 2]`/`r coinf_within_inf[7, 3]`, `r coinf_within_inf[7, 4]`%),  and Chiroptera (`r coinf_within_inf[3, 2]`/`r coinf_within_inf[3, 3]`, `r coinf_within_inf[3, 4]`%). 

### Triple infected animals
```{r triples}
pos_final %>% 
  dplyr::filter(coinfect_status == "coinfection") %>% 
  group_by(predict_sample_id) %>% 
  dplyr::filter(n()>2) %>% 
  dplyr::select(predict_sample_id, country, 
                scientific_name_to_lowest_known_rank, viral_family, virus) %>% 
    arrange(predict_sample_id, viral_family, virus) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_material()
```

### GLM results-all
```{r model results all}
summary(model_list[[1]])

sjPlot::tab_model(model_list[[1]])
```

### GLM results-bats
```{r model results bats}
summary(model_list[[2]])

sjPlot::tab_model(model_list[[2]])
```

### GLM results-rodents
```{r model results rodents}
summary(model_list[[3]])

sjPlot::tab_model(model_list[[3]])
```

### Binomial tests-all
```{r binomial tests-all}
binom_tests_vir_fam %>% 
  dplyr::select(virus_pair, exp_prev, obs_prev, greater_than_random, 
                p_adj, signif) %>% 
  mutate(across(c(exp_prev, obs_prev, p_adj), 
                ~formatC(.x, format = "e", digits = 2))) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_material()
```


### Binomial tests-bats
```{r binomial tests-bats}
binom_tests_vir_fam_bats %>% 
  dplyr::select(virus_pair, exp_prev, obs_prev, greater_than_random, 
                p_adj, signif) %>% 
  mutate(across(c(exp_prev, obs_prev, p_adj), 
                ~formatC(.x, format = "e", digits = 2))) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_material()
```


### Binomial tests-rodents
```{r binomial tests-rodents}
binom_tests_vir_fam_rodents %>% 
  dplyr::select(virus_pair, exp_prev, obs_prev, greater_than_random, 
                p_adj, signif) %>% 
  mutate(across(c(exp_prev, obs_prev, p_adj), 
                ~formatC(.x, format = "e", digits = 2))) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_material()
```

### Pteropus and Rattus
```{r pteropus and rattus, echo = T}
# within the model data, look at coinfection by host genus and geographic region
# most of the genera with high numbers of coinfected animals are in SES Asia
model_data_all %>% 
  group_by(host_genus, geo_region, virus_coinf, .drop = F) %>% 
  count() %>% 
  arrange(desc(virus_coinf), desc(n)) 

# what was the geographic distribution of pteropus and rattus overall?
pcr_all %>% 
  dplyr::filter(host_genus %in% c("Pteropus", "Rattus")) %>%   
  distinct(predict_sample_id, .keep_all = T) %>% 
  group_by(host_genus, geo_region) %>% 
  count()

# look at coinfection rates for pteropus and rattus by region
model_data_all %>% 
  dplyr::filter(host_genus %in% c("Pteropus", "Rattus")) %>% 
  group_by(host_genus, geo_region, virus_coinf) %>% 
  count()
```

### Wild animals in captivity
```{r wild captive, echo = T}
model_data_all %>% 
  dplyr::filter(captivity_status == "wild animal in captivity") %>% 
  group_by(host_sci_name) %>% 
  count()
```
