---
title: "Useful summary calculations"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

tar_load(pcr_all)
tar_load(pcr_pos)
tar_load(coinf_df)
tar_load(pos_final)
tar_load(model_data_all)
tar_load(model_list)

tar_load(binom_tests_vir_fam)
tar_load(binom_tests_vir_fam_bats)
tar_load(binom_tests_vir_fam_rodents)

library(tidyverse)
library(knitr)
library(kableExtra)
```

## Results

```{r data subsets, echo=TRUE}

# unique specimens
unique_specimens <- pcr_all %>% 
  distinct(specimen_id, .keep_all = TRUE) 

# unique animals
unique_animals <-  pcr_all %>% 
  distinct(predict_sample_id, .keep_all = TRUE) 
```

```{r summary values, echo=TRUE}
# number of unique specimens
n_unique_specimens <- nrow(unique_specimens)
n_unique_specimens

# number of unique specimens by country
spec_by_country <- unique_specimens %>% 
  group_by(country) %>% 
  dplyr::summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  mutate(perc = n/n_unique_specimens*100)
spec_by_country

# countries with 10k + specimens
countries_10k <- spec_by_country %>% 
  dplyr::filter(n > 10000) %>% 
  select(country)

# number of unique specimens by taxa group
spec_by_taxa <- unique_specimens %>% 
  group_by(taxa_group) %>% 
  dplyr::summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  mutate(perc = round(n/n_unique_specimens*100, 1))
spec_by_taxa

# number of PCR tests performed
n_tests <- nrow(pcr_all)
n_tests

# number of positive PCR tests
n_pos_tests <- nrow(pcr_pos)
n_pos_tests

# number of PCR tests by test type
table(pcr_all$viral_family_tested)
  
# number of unique animals
n_unique_animals <- nrow(unique_animals)
n_unique_animals

# number of unique animals by taxa group
table(unique_animals$taxa_group)

# number of infected animals
n_inf_animals <- nrow(coinf_df)
n_inf_animals

# overall prevalence (number animals pos. for 1+ virus/number of unique animals)
overall_prev <- round(n_inf_animals/n_unique_animals*100, 2)

# infected animals by country
coinf_df %>% 
  group_by(country) %>% 
  dplyr::summarise(n = n()) %>% 
  arrange(desc(n)) 

# number of coinfected animals
n_coinf_animals <- coinf_df %>% 
  dplyr::filter(n_virus > 1) %>% 
  nrow()
n_coinf_animals

# frequency of single/double/triple infection
coinf_freq <- table(coinf_df$n_virus)

# as percentages
coinf_perc <- round(coinf_freq/n_inf_animals*100,2)

# within COINFECTED animals, what was the breakdown by taxonomic group?
coinf_taxa_summary <- coinf_df %>% 
  dplyr::filter(n_virus > 1) %>% 
  group_by(taxa_group) %>% 
  summarize(n = n()) %>% 
  mutate(perc = round(n / sum(n)*100, 1))
coinf_taxa_summary

# within ALL infected animals, how common is coinfection by taxa group?
coinf_within_inf <- xtabs(~coinf_df$taxa_group + coinf_df$virus_coinf) %>% 
  as.data.frame.matrix() %>% 
  rename(single_inf = "0", coinf = "1") %>% 
  mutate(n_tot = single_inf + coinf,
         prev = round(coinf/n_tot*100, 2))

# prevalence by taxa group (within unique animals) for taxa groups with at least one infected animal
taxa_prev <- table(coinf_df$taxa_group) %>% 
  data.frame() %>% 
  rename(taxa_group = Var1, n_inf = Freq) %>% 
  left_join(
    table(unique_animals$taxa_group) %>% 
      data.frame() %>% 
      rename(taxa_group = Var1, n_tot = Freq)) %>% 
  mutate(prev = round(n_inf/n_tot*100, 2))

# coinfection by taxa group within the model data (subset of animals with age and sex)
model_coinf_within_inf <- xtabs(~taxa_group + virus_coinf, model_data_all) %>% 
  as.data.frame.matrix() %>% 
  rename(single_inf = "0", coinf = "1") %>% 
  mutate(n_tot = single_inf + coinf,
         prev = round(coinf/n_tot*100, 2))

```

### Methods (PREDICT-2 data overview)
Data from the PREDICT-2 project (2015-2019; the second half of the decade-long initiative) were openly shared on the USAID Development Data Library prior to the takedown of the USAID website. Specimen collection (n = `r n_unique_specimens` unique specimens from `r n_unique_animals` unique animals) was concentrated in countries in West and Central Africa as well as South and Southeast Asia, with > 10,000 specimens collected in `r countries_10k[1, 1]`, `r countries_10k[2, 1]`, and `r countries_10k[3, 1]` each (Fig. S1). Most specimens were collected from bats (n = `r spec_by_taxa[1,2]`, `r spec_by_taxa[1,3]`%) and rodents (n = `r spec_by_taxa[2,2]`, `r spec_by_taxa[2,3]`%). The dataset included some farm and companion animals, including poultry, dogs, and swine. Virus screening was conducted using consensus polymerase chain reaction (PCR) assays (`r n_tests`) total tests), mostly targeting viral families of significance to human health (e.g., coronaviruses, filoviruses, and paramyxoviruses), but in some cases testing for specific viruses of interest (e.g., influenza viruses and ebolaviruses). Further details on PREDICT can be found in the PREDICT Legacy Book (https://ohi.vetmed.ucdavis.edu/programs-projects/predict-project/reports).

### Results (overview)
We found that both infection and coinfection were rare overall. Out of `r n_unique_animals` unique animals, `r n_inf_animals` (`r overall_prev`%) were infected with one or more viruses (Fig. 1A); viral prevalence in birds (`r taxa_prev[2,2]`/`r taxa_prev[2,3]`; `r taxa_prev[2,4]`%) and bats (`r taxa_prev[1,2]`/`r taxa_prev[1,3]`; `r taxa_prev[1,4]`%) was somewhat higher than average, while prevalence in rodents (`r taxa_prev[8,2]`/`r taxa_prev[8,3]`; `r taxa_prev[8,4]`%) was slightly lower than average (Table S1).

A single virus was detected in most infected animals (n = `r coinf_freq[1]`, `r coinf_perc[1]`%); coinfections of two (n = `r coinf_freq[2]`, `r coinf_perc[2]`%) or three (n = `r coinf_freq[3]`, `r coinf_perc[3]`%) viruses in the same animal were infrequent. Of the `r n_coinf_animals` animals in which virus coinfection was detected, over half were bats (n = `r coinf_taxa_summary[1,2]`, `r coinf_taxa_summary[1,3]`%) with the remainder being `r coinf_taxa_summary[3,2]` rodents, `r coinf_taxa_summary[2,2]` birds, `r coinf_taxa_summary[5,2]` swine, and `r coinf_taxa_summary[4,2]` shrew. Within infected animals, virus coinfection rates were highest in rodents (`r coinf_within_inf[8, 2]`/`r coinf_within_inf[8, 3]`, `r coinf_within_inf[8, 4]`%), birds (`r coinf_within_inf[2, 2]`/`r coinf_within_inf[2, 3]`, `r coinf_within_inf[2, 4]`%), and bats (`r coinf_within_inf[1, 2]`/`r coinf_within_inf[1, 3]`, `r coinf_within_inf[1, 4]`%). 

### Triple infected animals
```{r triples}
pos_final %>% 
  dplyr::filter(coinfect_status == "coinfection") %>% 
  group_by(predict_sample_id) %>% 
  dplyr::filter(n()>2) %>% 
  dplyr::select(predict_sample_id, country, taxa_group, 
                scientific_name_to_lowest_known_rank, viral_family, virus) %>% 
    arrange(taxa_group, predict_sample_id, viral_family, virus) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_material()
```

### GLM results-all
```{r model results all}
summary(model_list[[1]])

sjPlot::tab_model(model_list[[1]])
```

### GLM results-bats
```{r model results bats}
summary(model_list[[2]])

sjPlot::tab_model(model_list[[2]])
```

### GLM results-rodents
```{r model results rodents}
summary(model_list[[3]])

sjPlot::tab_model(model_list[[3]])
```

### Binomial tests-all
```{r binomial tests-all}
binom_tests_vir_fam %>% 
  dplyr::select(virus_pair, exp_prev, obs_prev, greater_than_random, 
                p_adj, signif) %>% 
  mutate(across(c(exp_prev, obs_prev, p_adj), 
                ~formatC(.x, format = "e", digits = 2))) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_material()
```


### Binomial tests-bats
```{r binomial tests-bats}
binom_tests_vir_fam_bats %>% 
  dplyr::select(virus_pair, exp_prev, obs_prev, greater_than_random, 
                p_adj, signif) %>% 
  mutate(across(c(exp_prev, obs_prev, p_adj), 
                ~formatC(.x, format = "e", digits = 2))) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_material()
```


### Binomial tests-rodents
```{r binomial tests-rodents}
binom_tests_vir_fam_rodents %>% 
  dplyr::select(virus_pair, exp_prev, obs_prev, greater_than_random, 
                p_adj, signif) %>% 
  mutate(across(c(exp_prev, obs_prev, p_adj), 
                ~formatC(.x, format = "e", digits = 2))) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_material()
```

### Pteropus and Rattus
```{r pteropus and rattus, echo = T}
# within the model data, look at coinfection by host genus and geographic region
# most of the genera with high numbers of coinfected animals are in SES Asia
model_data_all %>% 
  group_by(host_genus, geo_region, virus_coinf, .drop = F) %>% 
  count() %>% 
  arrange(desc(virus_coinf), desc(n)) 

# what was the geographic distribution of pteropus and rattus overall?
pcr_all %>% 
  dplyr::filter(host_genus %in% c("Pteropus", "Rattus")) %>%   
  distinct(predict_sample_id, .keep_all = T) %>% 
  group_by(host_genus, geo_region) %>% 
  count()

# look at coinfection rates for pteropus and rattus by region
model_data_all %>% 
  dplyr::filter(host_genus %in% c("Pteropus", "Rattus")) %>% 
  group_by(host_genus, geo_region, virus_coinf) %>% 
  count()
```

### Wild animals in captivity
```{r wild captive, echo = T}
model_data_all %>% 
  dplyr::filter(captivity_status == "wild animal in captivity") %>% 
  group_by(host_sci_name) %>% 
  count()
```
